---
title: "Intro to Time Series Regression"
author: "William Chiu"
date: "2/23/2022"
output:
  powerpoint_presentation:
    reference_doc: william_template.pptx
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

options(digits=3)

library(tidyverse)
library(broom)
library(broom.mixed)
library(np)
library(nlme)
library(lmtest)
library(sandwich)
library(purrr)
library(furrr)
library(ModelMetrics)

future::plan(multisession, workers=3)

Ksim <- 1000
boot_samp <- 1000
```

## About Me

- MSIA Class of 2013
- Works for a bank
- Team develops models to explain and forecast:
  - Credit Losses
  - Loan Prepayments
  - Mortgage Rates
  - Deposit Attrition
- Primarily uses SQL and R


## Time Series Regression

* Time series is data collected in equally-spaced time intervals.

* Linear regression estimates the linear relationship between a continuous response (`y`) and one or more predictors (`x`).

* Ordinary least squares (OLS) is the most common implementation of linear regression, and estimates the coefficients that minimize the error sum of squares:

$ESS = \sum_{i=1}^{n}{\left ( y_i - (\widehat{\beta_0} + \widehat{\beta_1}X_{1,i} + ... + \widehat{\beta_p}X_{p,i}) \right )^2}$

* Using linear regression on time series is called time series regression.




## OLS Assumptions for Time Series

* Population data generating process (dgp) most suitable for OLS: 

<br>

$y_t = \beta_0 + \beta_1X_{1,t} + ... + \beta_pX_{p,t} + \varepsilon_t$

$\varepsilon_t \sim \mathcal{N}(0, \sigma^2)$

<br>

* Linear parameters

* Independent errors (no serial correlation)

* Normally distributed errors

* Equal variance errors (homoscedasticity)

* Stationary errors

## Violation: Serial Correlation

* Unfortunately, the data generating process (dgp) for time series often violates the independent errors assumption. For example:

<br>

$y_t = \beta_0 + \beta_1X_{1,t} + ... + \beta_pX_{p,t} + \varepsilon_t$

$\varepsilon_t = \rho\varepsilon_{t-1} + v_t$

$v_t \sim \mathcal{N}(0, \sigma^2)$

<br>

* If $\rho=0$, then there is no serial correlation. 

* If $-1<\rho<1$, then there is serial correlation. 

## Simulate predictors with serial correlation

```{r, echo=TRUE}
set.seed(2013) 

x1 <- arima.sim(list(order = c(1,0,0), ar = 0.6), n = 200, sd=2, mean=2)
x2 <- arima.sim(list(order = c(1,0,0), ar = -0.6), n = 200, sd=3, mean=2)
x3 <- arima.sim(list(order = c(1,0,0), ar = 0.6), n = 200, sd=4, mean=2)
```

## Simulate response with serial correlation

```{r, echo=TRUE}
epsilon <- arima.sim(list(order = c(1,0,0), ar = 0.6), n = 200, sd=5, mean=0)

y <- 10 + 3.5*x1 - 1.5*x2 + epsilon 
## x3 is not part of the dgp

df <- data.frame(y=y, x1=x1, x2=x2, x3=x3, t=1:length(y))
```

## Plot simulated data (first 50 obs)

```{r, warning=FALSE, message=FALSE, fig.width = 14,fig.height = 7}

cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#000000", "#F0E442",
                "#0072B2", "#D55E00", "#CC79A7")

df_plot <- df %>% 
  slice_head(n=50) %>% # first 50 obs
  pivot_longer(-t, "variable", "value") %>% 
  arrange(desc(variable), t) # preserve time order

ggplot(df_plot, aes(x=t, y=value, color=variable)) +
  geom_line(size=2) +
  theme_bw(base_size=28) +
  scale_colour_manual(values=cbbPalette)

```

## Split between train and test

```{r, echo=TRUE}
train <- head(df, 150)
test <- tail(df, 50)
```

## OLS ignores serial correlation

:::::::::::::: {.columns}
::: {.column}
* Under serial correlation, OLS parameter estimates are unbiased and consistent
* However, standard errors are incorrect
* And p-values are incorrect
* `x3` should *not* be statistically significant
:::
::: {.column}
```{r}

lm_mod <- lm(y ~ x1 + x2 + x3, data=train)

knitr::kable(tidy(lm_mod))
```
:::
::::::::::::::

## Detecting serial correlation

:::::::::::::: {.columns}
::: {.column}
* Durbin-Watson test is a popular test for serial correlation, but suffers from some major drawbacks. 

* Breusch-Godfrey Test is a better test.
:::
::: {.column}
```{r}
knitr::kable(tidy(bgtest(lm_mod)), digits=28, caption="Null Hypothesis: No serial correlation")
```
:::
::::::::::::::


## Remediating serial correlation

* Replace OLS with GLS
* Keep OLS but estimate robust (HAC) standard errors
* Keep OLS but bootstrap the standard errors

## Generalized Least Squares (Cochrane–Orcutt)

:::::::::::::: {.columns}
::: {.column}
1. Estimate $\rho$
2. Transform each response and predictor into "partial differences" (e.g., $y_t - \widehat{\rho}y_{t-1}$)
3. Estimate the transformed model using OLS
4. Re-calculate the coefficient estimates in terms of the original training data
:::
::: {.column}
```{r, message=FALSE, warning=FALSE}

gls_mod <- gls(y ~ x1 + x2 + x3, data=train, correlation=corARMA(p=1, q=0), method="ML")

knitr::kable(tidy(gls_mod))
```
:::
::::::::::::::






## Newey-West HAC Correction

:::::::::::::: {.columns}
::: {.column}
* Keep OLS but estimate robust (HAC) standard errors

:::
::: {.column}
```{r, message=FALSE, warning=FALSE}
nw <- coeftest(lm_mod, vcov=NeweyWest)

knitr::kable(tidy(nw ## Newey-West correction
))
```
:::
::::::::::::::





## Andrews HAC Correction

:::::::::::::: {.columns}
::: {.column}
* Keep OLS but estimate robust (HAC) standard errors

:::
::: {.column}
```{r, message=FALSE, warning=FALSE}
andrews <- coeftest(lm_mod, vcov=kernHAC)

knitr::kable(tidy(andrews ## Andrews correction
))

```
:::
::::::::::::::


## Block Bootstrap


:::::::::::::: {.columns}
::: {.column}
* Keep the OLS coefficient estimates but *bootstrap* the standard errors
* Break the time series into sequential blocks (non-random)
* Randomly sample blocks with replacement
* Unfortunately, results are sensitive to block size

:::
::: {.column}
```{r, message=FALSE, warning=FALSE}
even_blocksize <- which(nrow(train) %% c(1:nrow(train))==0)

optimal_blocksize <- max(b.star(lm_mod$residuals))
    
optimal_blocksize <- even_blocksize[which.min(abs(even_blocksize - optimal_blocksize))]

nbrClusters <- nrow(train)/optimal_blocksize

clusterID <- rep(seq(nbrClusters),each=optimal_blocksize)

set.seed(1)

blockboot <- coeftest(lm_mod, vcov=vcovBS, cluster=clusterID, R=boot_samp)
knitr::kable(tidy(blockboot
))
```
:::
::::::::::::::

## Test Set Performance (Ex-post, OLS vs. GLS)

```{r, fig.width = 14,fig.height = 7}
ols_pred <- predict(lm_mod, newdata=test)
gls_pred <- predict(gls_mod, newdata=test)

test_plot <- test

test_plot$ols <- ols_pred
test_plot$gls <- gls_pred

test_plot <- test_plot %>% 
  select(t, y, ols, gls) %>% 
  pivot_longer(-t, "variable", "value") %>% 
  arrange(variable, t)

cbbPalette2 <- c("#E69F00", "#56B4E9", "#000000", "#009E73", "#F0E442",
                 "#0072B2", "#D55E00", "#CC79A7")

ggplot(test_plot, aes(x=t, y=value, color=variable, lty=variable)) +
  geom_line(size=1.5) +
  theme_bw(base_size=28) +
  scale_colour_manual(values=cbbPalette2) +
  scale_linetype_manual(values=c("dotted","dashed","solid"))
```


## Inference is harder than prediction

* Test set performance appears similar between OLS and GLS

* However, the p-values for `x3` differ substantially by method: `lm`, `gls`, `newey-west`, `andrews`, `bootstrap`.

* The p-values from simulated data are *sensitive* to the random seeds

* In order to measure the quality of the methods: 

   1. Re-simulate the training data ($K = `r Ksim`$) with varying seeds and $\rho$
   2. Measure how often Type 1 errors occur ($\alpha=0.10$)
   3. Measure test set error

* Since $\beta_3 = 0$ in the simulation, any p-values <= $\alpha$ is a Type 1 Error

## Type 1 Error Frequency

:::::::::::::: {.columns}
::: {.column}
```{r, message=FALSE, warning=FALSE}
doOneSimulation <- function(randomSeed=2013, alpha=0.10, rho=0.6) {
  set.seed(randomSeed) 
  
  x1 <- arima.sim(list(order = c(1,0,0), ar = rho), n = 200, sd=2, mean=2)
  x2 <- arima.sim(list(order = c(1,0,0), ar = rho*-1), n = 200, sd=3, mean=2)
  x3 <- arima.sim(list(order = c(1,0,0), ar = rho), n = 200, sd=4, mean=2)
  
  epsilon <- arima.sim(list(order = c(1,0,0), ar = rho), n = 200, sd=5, mean=0)
  
  y <- 10 + 3.5*x1 - 1.5*x2 + epsilon 
  ## x3 is not part of the dgp
  
  df <- data.frame(y=y, x1=x1, x2=x2, x3=x3, t=1:length(y))
  
  train <- head(df, 150)
  test <- tail(df, 50)
  
  lm_mod <- lm(y ~ x1 + x2 + x3, data=train)
  
  gls_mod <- gls(y ~ x1 + x2 + x3, data=train, correlation=corARMA(p=1, q=0), method="ML")
  
  nw <- coeftest(lm_mod, vcov=NeweyWest)
  
  andrews <- coeftest(lm_mod, vcov=kernHAC)
  
  even_blocksize <- which(nrow(train) %% c(1:nrow(train))==0)
  
  optimal_blocksize <- max(b.star(lm_mod$residuals))
      
  optimal_blocksize <- even_blocksize[which.min(abs(even_blocksize - optimal_blocksize))]
  
  nbrClusters <- nrow(train)/optimal_blocksize
  
  clusterID <- rep(seq(nbrClusters),each=optimal_blocksize)
  
  set.seed(1)
  
  blockboot <- coeftest(lm_mod, vcov=vcovBS, cluster=clusterID, R=boot_samp)
  
  all_models <- list(lm=lm_mod, gls=gls_mod, nw=nw, andrews=andrews, bootstrap=blockboot)
  
  df_of_models <- map_dfr(all_models, tidy, .id="method")
  
  out <- df_of_models %>% 
    filter(term == "x3") %>% 
    mutate(Type1ErrorFlag = `p.value` <= alpha)
  
  ## test set performance
  
  ols_pred <- predict(lm_mod, newdata=test)
  gls_pred <- predict(gls_mod, newdata=test)
  
  test_plot <- test
  
  test_plot$ols <- ols_pred
  test_plot$gls <- gls_pred
  
  test_rmse_lm <- data.frame(method='lm', Test_RMSE = rmse(actual=test_plot$y,
                                                           predicted=test_plot$ols))
  test_rmse_gls <- data.frame(method='gls', Test_RMSE = rmse(actual=test_plot$y,
                                                             predicted=test_plot$gls))
  
  test_rmse <- bind_rows(test_rmse_lm, test_rmse_gls)
  
  ## merge results
  
  out <- out %>% 
    left_join(test_rmse, by="method")

  return(out)
}

a <- Sys.time()

allSimulations <- map_dfr(c(`1-Low (0.1)`=0.1, `2-High (0.7)`=0.7, `3-Very High (0.99)`=0.99),
                          ~ future_map_dfr(1:Ksim, doOneSimulation, rho=.x, .options =
                                             furrr_options(packages=c("broom", "broom.mixed"),
                                                           seed=NULL)),
                          .id="rho")

b <- Sys.time()
c <- difftime(b, a, units="mins")

future:::ClusterRegistry("stop")

tbl <- allSimulations %>% 
  group_by(rho, method) %>% 
  summarize(Prob_Type_1_Error = mean(Type1ErrorFlag),
            Avg_Test_RMSE = mean(Test_RMSE),
            .groups = 'drop') %>% 
  arrange(rho, Prob_Type_1_Error) 

knitr::kable(head(tbl,5), digits=3)
```

:::
::: {.column}

```{r}
knitr::kable(tbl[6:10,], digits=3)
```
:::
::::::::::::::







## Very High Serial Correlation

:::::::::::::: {.columns}
::: {.column}

* Under very high serial correlation, most of the previous methods struggle to return the correct p-values.

* If $\widehat{\rho}$ is close to -1 or 1, then check for non-stationary errors (e.g., Phillips-Ouliaris Cointegration Test).

:::
::: {.column}
```{r}
knitr::kable(tail(tbl,5), digits=3)

```
:::
::::::::::::::


## Violation: Non-stationary errors

* When $\rho=1$ or $\rho=-1$, the errors are non-stationary.

<br>

$\varepsilon_t = \rho\varepsilon_{t-1} + v_t$

<br>

* Under non-stationary errors, OLS and GLS are both biased and inconsistent (i.e., spurious)

* Non-stationary errors adversely affect *both* prediction and inference (while serially correlated errors affect only inference)

## Remediating non-stationary errors

1. Transform each response and predictor into "first differences" (e.g., $y_t - y_{t-1}$)
2. Estimate the transformed model using OLS or GLS
3. Forecast with the last historical value and the cumulative sum of the predicted first differences:
  
$\widehat{y_{t+h}} = y_t + \sum_{i=t+1}^{t+h}{\widehat{\Delta{y_i}}}$

## Conclusions

* OLS is most suitable for independent errors

* The independent error assumption is often violated with time series regression

* Under serial correlation, standard errors and p-values from OLS are unreliable

* Fortunately, OLS coefficient estimates remain unbiased and consistent

* Remediating serial correlation:
  - GLS
  - OLS with HAC standard errors
  - OLS with block bootstrapped standard errors
  
* Under non-stationary errors, beware of spurious regressions



